---
title: "Codebook for ECON490 Final Paper"
author: "Joanna Wang, 54536388"
date: "2023-10-09"
output: html_document
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load Packages

```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(readxl)
```

# Preprocessing

```{r}
years <- 2015:2022
data_list <- list()

for(year in years) {
  data_list[[paste0('w',year)]] <- read.csv(paste0("~/Desktop/ECON490/WHR/data/", year, ".csv"))
}

region <- read.csv('~/Desktop/ECON490/WHR/data/region.csv')
income <- readxl::read_xlsx('~/Desktop/ECON490/WHR/data/income.xlsx')
pop <- read.csv('~/Desktop/ECON490/WHR/data/population.csv')
gdp <- read.csv('~/Desktop/ECON490/WHR/data/gdp.csv')
gdp_22 <- readxl::read_xls('~/Desktop/ECON490/WHR/data/gdp_2022.xls')
life <- read.csv('~/Desktop/ECON490/WHR/data/life.csv')
life_22 <- read.csv('~/Desktop/ECON490/WHR/data/life_2022.csv')
medage <- read.csv('~/Desktop/ECON490/WHR/data/median-age.csv')
medage_22 <- readxl::read_xlsx('~/Desktop/ECON490/WHR/data/median-age_22.xlsx')
list2env(data_list, envir = .GlobalEnv)
```

## ETL

```{r}
gdp <- gdp %>%
  rename(country = Entity, year = Year, gdp = GDP.per.capita..PPP..constant.2017.international...) %>%
  select(country, year, gdp) %>%
  bind_rows(gdp_22) %>%
  mutate(gdp = ifelse(gdp != 0, log(gdp), 0)) %>%
  rename(gdp_ppp = gdp) 

life_exp <- life %>%
  rename(country = Entity, year = Year, life = Life.expectancy.at.birth..historical.) %>%
  select(country, year, life) %>%
  bind_rows(life_22)
```

```{r}
whr_2015 <- w2015 %>%
  select(
    country = grep("country", names(w2015), ignore.case = TRUE, value = TRUE)[1],
    rank = grep("rank", names(w2015), ignore.case = TRUE, value = TRUE)[1],
    score = grep("score", names(w2015), ignore.case = TRUE, value = TRUE)[1],
    gdp = grep("GDP.per.capita", names(w2015), ignore.case = TRUE, value = TRUE)[1],
    social_support = grep("family|social.support", names(w2015), ignore.case = TRUE, value = TRUE)[1],
    health = grep("health", names(w2015), ignore.case = TRUE, value = TRUE)[1],
    freedom = grep("freedom", names(w2015), ignore.case = TRUE, value = TRUE)[1],
    corruption = grep("corruption", names(w2015), ignore.case = TRUE, value = TRUE)[1],
    generosity = grep("generosity", names(w2015), ignore.case = TRUE, value = TRUE)[1]
  ) %>%
  mutate(year = 2015) %>%
  left_join(gdp, by = c("country", "year")) %>%
  left_join(life_exp, by = c("country", "year"))

whr_2016 <- w2016 %>%
  select(
    country = grep("country", names(w2016), ignore.case = TRUE, value = TRUE)[1],
    rank = grep("rank", names(w2016), ignore.case = TRUE, value = TRUE)[1],
    score = grep("score", names(w2016), ignore.case = TRUE, value = TRUE)[1],
    gdp = grep("GDP.per.capita", names(w2016), ignore.case = TRUE, value = TRUE)[1],
    social_support = grep("family|social.support", names(w2016), ignore.case = TRUE, value = TRUE)[1],
    health = grep("health", names(w2016), ignore.case = TRUE, value = TRUE)[1],
    freedom = grep("freedom", names(w2016), ignore.case = TRUE, value = TRUE)[1],
    corruption = grep("corruption", names(w2016), ignore.case = TRUE, value = TRUE)[1],
    generosity = grep("generosity", names(w2016), ignore.case = TRUE, value = TRUE)[1]
  )%>%
  mutate(year = 2016)%>%
  left_join(gdp, by = c("country", "year")) %>%
  left_join(life_exp, by = c("country", "year"))

whr_2017 <- w2017 %>%
  select(
    country = grep("country", names(w2017), ignore.case = TRUE, value = TRUE)[1],
    rank = grep("rank", names(w2017), ignore.case = TRUE, value = TRUE)[1],
    score = grep("score", names(w2017), ignore.case = TRUE, value = TRUE)[1],
    gdp = grep("GDP.per.capita", names(w2017), ignore.case = TRUE, value = TRUE)[1],
    social_support = grep("family|social.support", names(w2017), ignore.case = TRUE, value = TRUE)[1],
    health = grep("health", names(w2017), ignore.case = TRUE, value = TRUE)[1],
    freedom = grep("freedom", names(w2017), ignore.case = TRUE, value = TRUE)[1],
    corruption = grep("corruption", names(w2017), ignore.case = TRUE, value = TRUE)[1],
    generosity = grep("generosity", names(w2017), ignore.case = TRUE, value = TRUE)[1]
  )%>%
  mutate(year = 2017)%>%
  left_join(gdp, by = c("country", "year")) %>%
  left_join(life_exp, by = c("country", "year"))

whr_2018 <- w2018 %>%
  select(
    country = grep("country", names(w2018), ignore.case = TRUE, value = TRUE)[1],
    rank = grep("rank", names(w2018), ignore.case = TRUE, value = TRUE)[1],
    score = grep("score", names(w2018), ignore.case = TRUE, value = TRUE)[1],
    gdp = grep("GDP.per.capita", names(w2018), ignore.case = TRUE, value = TRUE)[1],
    social_support = grep("family|social.support", names(w2018), ignore.case = TRUE, value = TRUE)[1],
    health = grep("health", names(w2018), ignore.case = TRUE, value = TRUE)[1],
    freedom = grep("freedom", names(w2018), ignore.case = TRUE, value = TRUE)[1],
    corruption = grep("corruption", names(w2018), ignore.case = TRUE, value = TRUE)[1],
    generosity = grep("generosity", names(w2018), ignore.case = TRUE, value = TRUE)[1]
  )%>%
  mutate(corruption = as.numeric(corruption)) %>%
  mutate(year = 2018)%>%
  left_join(gdp, by = c("country", "year")) %>%
  left_join(life_exp, by = c("country", "year"))

whr_2019 <- w2019 %>%
  select(
    country = grep("country", names(w2019), ignore.case = TRUE, value = TRUE)[1],
    rank = grep("rank", names(w2019), ignore.case = TRUE, value = TRUE)[1],
    score = grep("score", names(w2019), ignore.case = TRUE, value = TRUE)[1],
    gdp = grep("GDP.per.capita", names(w2019), ignore.case = TRUE, value = TRUE)[1],
    social_support = grep("family|social.support", names(w2019), ignore.case = TRUE, value = TRUE)[1],
    health = grep("health", names(w2019), ignore.case = TRUE, value = TRUE)[1],
    freedom = grep("freedom", names(w2019), ignore.case = TRUE, value = TRUE)[1],
    corruption = grep("corruption", names(w2019), ignore.case = TRUE, value = TRUE)[1],
    generosity = grep("generosity", names(w2019), ignore.case = TRUE, value = TRUE)[1]
  )%>%
  mutate(year = 2019)%>%
  left_join(gdp, by = c("country", "year")) %>%
  left_join(life_exp, by = c("country", "year"))

whr_2020 <- w2020 %>%
  select(
    country = grep("country", names(w2020), ignore.case = TRUE, value = TRUE)[1],
    score = grep("score", names(w2020), ignore.case = TRUE, value = TRUE)[1],
    gdp = grep("Explained.by..Log.GDP", names(w2020), ignore.case = TRUE, value = TRUE)[1],
    social_support = grep("Explained.by..Social", names(w2020), ignore.case = TRUE, value = TRUE)[1],
    health = grep("Explained.by..Healthy", names(w2020), ignore.case = TRUE, value = TRUE)[1],
    freedom = grep("Explained.by..Freedom", names(w2020), ignore.case = TRUE, value = TRUE)[1],
    corruption = grep("Explained.by..Perceptions", names(w2020), ignore.case = TRUE, value = TRUE)[1],
    generosity = grep("Explained.by..Generosity", names(w2020), ignore.case = TRUE, value = TRUE)[1],
    gdp_ppp = grep("Logged.GDP.per.capita", names(w2020), ignore.case = TRUE, value = TRUE)[1],
    life = grep("Healthy.life.expectancy", names(w2020), ignore.case = TRUE, value = TRUE)[1]
  )%>%
  mutate(rank = rank(-score, ties.method = "min"))%>%
  mutate(year = 2020)

whr_2021 <- w2021 %>%
  select(
    country = grep("country", names(w2021), ignore.case = TRUE, value = TRUE)[1],
    score = grep("score", names(w2021), ignore.case = TRUE, value = TRUE)[1],
    gdp = grep("Explained.by..Log.GDP", names(w2021), ignore.case = TRUE, value = TRUE)[1],
    social_support = grep("Explained.by..Social", names(w2021), ignore.case = TRUE, value = TRUE)[1],
    health = grep("Explained.by..Healthy", names(w2021), ignore.case = TRUE, value = TRUE)[1],
    freedom = grep("Explained.by..Freedom", names(w2021), ignore.case = TRUE, value = TRUE)[1],
    corruption = grep("Explained.by..Perceptions", names(w2021), ignore.case = TRUE, value = TRUE)[1],
    generosity = grep("Explained.by..Generosity", names(w2021), ignore.case = TRUE, value = TRUE)[1],
    gdp_ppp = grep("Logged.GDP.per.capita", names(w2021), ignore.case = TRUE, value = TRUE)[1],
    life = grep("Healthy.life.expectancy", names(w2021), ignore.case = TRUE, value = TRUE)[1]
  )%>%
  mutate(rank = rank(-score, ties.method = "min"))%>%
  mutate(year = 2021)

whr_2022 <- w2022 %>%
  select(
    country = grep("country", names(w2022), ignore.case = TRUE, value = TRUE)[1],
    score = grep("score", names(w2022), ignore.case = TRUE, value = TRUE)[1],
    gdp = grep("Explained.by..GDP", names(w2022), ignore.case = TRUE, value = TRUE)[1],
    social_support = grep("Explained.by..Social", names(w2022), ignore.case = TRUE, value = TRUE)[1],
    health = grep("Explained.by..Healthy", names(w2022), ignore.case = TRUE, value = TRUE)[1],
    freedom = grep("Explained.by..Freedom", names(w2022), ignore.case = TRUE, value = TRUE)[1],
    corruption = grep("Explained.by..Perceptions", names(w2022), ignore.case = TRUE, value = TRUE)[1],
    generosity = grep("Explained.by..Generosity", names(w2022), ignore.case = TRUE, value = TRUE)[1]
  ) %>%
  mutate_at(vars(score, gdp, social_support, health, freedom, corruption, generosity),
            ~ as.numeric(gsub(",", ".", as.character(.)))) %>%
  mutate(rank = rank(-score, ties.method = "min"))%>%
  mutate(year = 2022)%>%
  left_join(gdp, by = c("country", "year")) %>%
  left_join(life_exp, by = c("country", "year"))
```

```{r}
whr_df <- bind_rows(whr_2015, whr_2016, whr_2017, whr_2018, whr_2019, whr_2020, whr_2021, whr_2022)

colnames(whr_df)
```

## Handling NAs

```{r}
na_count <- sapply(whr_df, function(y) sum(length(which(is.na(y)))))
print(na_count)

rows_with_na <- which(rowSums(is.na(whr_df)) > 0)
print(rows_with_na)

whr_df_clean <- whr_df[!rowSums(is.na(whr_df)),]
```

## Select countries

```{r}
country_counts <- whr_df_clean %>%
                  group_by(country) %>%
                  summarise(year_count = n())

full_data_countries <- filter(country_counts, year_count == 8) %>%
                       pull(country)
whr <- filter(whr_df_clean, country %in% full_data_countries) %>%
  mutate(gdp_ppp = ifelse(country == "Afghanistan" & year == 2022, 5.8090730, gdp_ppp))

unique_countries <- unique(whr$country)
```

## Add IV Median Age

```{r}
medage_22 <- rename(medage_22, year = Year, country = region) %>%
  mutate(median_age = as.numeric(median_age))
medage <- select(medage, -Code) %>%
  rename(., year = Year, country = Entity)

medage_all <- bind_rows(medage, medage_22)


whr <- whr %>%
  left_join(medage_all, by=c("country","year"))

whr <- whr %>%
  arrange(country, year, is.na(median_age)) %>%
  group_by(country, year) %>%
  filter(!is.na(median_age) | row_number() == 1) %>%
  ungroup()


whr <- whr %>%
  mutate(median_age = case_when(
    country == "Bolivia" & year == 2022 & is.na(median_age) ~ 24.4,
    country == "Iran" & year == 2022 & is.na(median_age) ~ 33.0,
    country == "Moldova" & year == 2022 & is.na(median_age) ~ 34.7,
    country == "Tanzania" & year == 2022 & is.na(median_age) ~ 17.0,
    country == "United States" & year == 2022 & is.na(median_age) ~ 38.9,
    country == "Vietnam" & year == 2022 & is.na(median_age) ~ 32.8,
    TRUE ~ median_age
  ))
```
## Add other predictors
```{r}
indicators <- read.csv("/Users/mac/Desktop/ECON490/WHR/data/indicators2.csv")

whr <- whr %>%
  left_join(.,indicators, by=c("country","year")) %>%
  rename(unemp_rate = Unemployment.rate, population = Population.size.in.millions, pct_urban_population = Percent.urban.population)
```


# EDA

```{r warning=FALSE}
variables <- c("gdp", "social_support", "health", "freedom", "corruption", "generosity", "gdp_ppp", "life", "unemp_rate", "population", "pct_urban_population", "inflation")

for (var in variables) {
  
  model <- lm(paste("score ~", var), data = whr)
  coef_value <- coef(model)[2]
  
  print(
    ggplot(whr, aes_string(x = var, y = "score")) + 
      geom_point(aes_string(color = var)) + 
      geom_smooth(method = "lm", se = FALSE, color = "red") + 
      annotate("text", x = Inf, y = Inf, 
               label = paste("Coefficient =", round(coef_value, 4)), 
               hjust = 1, vjust = 1, size = 5, color = "red") +
      labs(title = paste("Scatterplot of", var, "VS Score"),
           x = var,
           y = "Score") +
      theme_minimal()
  )
}

for (var in variables) {
  p <- ggplot(whr, aes(x = as.factor(year), y = get(var))) + 
       geom_boxplot() +
       labs(title = paste("Boxplot of", var, "by Year"),
            x = "Year", 
            y = var) +
       theme_minimal()
  
  print(p)
}

```

# Event Study Framework

## Visualization

```{r}
data_2015 <- whr %>% filter(year == 2015)
data_2019 <- whr %>% filter(year == 2019)

# Compute the change in score
change_in_happiness <- data_2019 %>%
  select(country, score) %>%
  left_join(data_2015 %>% select(country, score_2015 = score), by = "country") %>%
  mutate(score_change = score - score_2015)

# Select the top and bottom countries based on the score change
top_bottom_n <- 10
top_countries <- head(change_in_happiness %>% arrange(desc(score_change)), n = top_bottom_n)
bottom_countries <- head(change_in_happiness %>% arrange(score_change), n = top_bottom_n)
plot_data <- rbind(top_countries, bottom_countries)

# Order the countries factor to match the order in the plot data
plot_data$country <- factor(plot_data$country, levels = rev(plot_data$country))

# Create a horizontal bar chart
p <- ggplot(plot_data, aes(x = country, y = score_change)) +
  geom_bar(stat = "identity", aes(fill = score_change > 0), position = position_dodge(width = 0.9)) +
  coord_flip() +  # Flip the axes to get a horizontal bar chart
  xlab('Country') +
  ylab('Change in Happiness Score') +
  ggtitle('Changes in Happiness from 2015 to 2019') +
  theme_minimal() +
  scale_fill_manual(values = c('skyblue', 'orange')) +
  geom_text(aes(label = sprintf('%.2f', score_change)), hjust = -0.1)

# Print the plot
print(p)
```

```{r}
data_2022 <- whr %>% filter(year == 2022)

# Compute the change in score
change_in_happiness <- data_2022 %>%
  select(country, score) %>%
  left_join(data_2019 %>% select(country, score_2019 = score), by = "country") %>%
  mutate(score_change = score - score_2019)

# Select the top and bottom countries based on the score change
top_bottom_n <- 10
top_countries <- head(change_in_happiness %>% arrange(desc(score_change)), n = top_bottom_n)
bottom_countries <- head(change_in_happiness %>% arrange(score_change), n = top_bottom_n)
plot_data <- rbind(top_countries, bottom_countries)

# Order the countries factor to match the order in the plot data
plot_data$country <- factor(plot_data$country, levels = rev(plot_data$country))

# Create a horizontal bar chart
p <- ggplot(plot_data, aes(x = country, y = score_change)) +
  geom_bar(stat = "identity", aes(fill = score_change > 0), position = position_dodge(width = 0.9)) +
  coord_flip() +  # Flip the axes to get a horizontal bar chart
  xlab('Country') +
  ylab('Change in Happiness Score') +
  ggtitle('Changes in Happiness from 2019 to 2022') +
  theme_minimal() +
  scale_fill_manual(values = c('skyblue', 'orange')) +
  geom_text(aes(label = sprintf('%.2f', score_change)), hjust = -0.1)

# Print the plot
print(p)
```

```{r}
# Compute the change in score
change_in_happiness <- data_2022 %>%
  select(country, score) %>%
  left_join(data_2015 %>% select(country, score_2015 = score), by = "country") %>%
  mutate(score_change = score - score_2015)

# Select the top and bottom countries based on the score change
top_bottom_n <- 10
top_countries <- head(change_in_happiness %>% arrange(desc(score_change)), n = top_bottom_n)
bottom_countries <- head(change_in_happiness %>% arrange(score_change), n = top_bottom_n)
plot_data <- rbind(top_countries, bottom_countries)

# Order the countries factor to match the order in the plot data
plot_data$country <- factor(plot_data$country, levels = rev(plot_data$country))

# Create a horizontal bar chart
p <- ggplot(plot_data, aes(x = country, y = score_change)) +
  geom_bar(stat = "identity", aes(fill = score_change > 0), position = position_dodge(width = 0.9)) +
  coord_flip() +  # Flip the axes to get a horizontal bar chart
  xlab('Country') +
  ylab('Change in Happiness Score') +
  ggtitle('Changes in Happiness from 2015 to 2022') +
  theme_minimal() +
  scale_fill_manual(values = c('skyblue', 'orange')) +
  geom_text(aes(label = sprintf('%.2f', score_change)), hjust = -0.1)

# Print the plot
print(p)

```

## Global Aggregation (Paired t test)
### Estimate the "Normal" Behavior

```{r}
normal_behavior <- whr %>%
                   filter(year %in% 2015:2019) %>%
                   group_by(country) %>%
                   summarise(avg_score_pre_covid = mean(score, na.rm=TRUE))
```

### Compute Abnormal Returns

```{r}
abnormal_returns <- whr %>%
                    filter(year %in% 2020:2022) %>%
                    left_join(normal_behavior, by="country") %>%
                    mutate(abnormal_score = score - avg_score_pre_covid) %>%
                    select(country, year, score, avg_score_pre_covid, abnormal_score)
```

### Test for Significance

```{r}
library(stats)
t.test(abnormal_returns$score, abnormal_returns$avg_score_pre_covid, paired=TRUE)
```


## Income Group Specific

```{r}
income <- income %>%
  rename(country = Economy, income_group = `Income group`) %>%
  select(country, income_group)

whr <- whr %>%
  left_join(income, by="country")

na_income_countries <- whr %>%
  filter(is.na(income_group)) %>%
  distinct(country) %>%
  pull(country)

whr <- whr %>%
  mutate(income_group = case_when(
    country %in% c("Slovakia", "South Korea") ~ "High income",
    country %in% c("Venezuela", "Russia", "Turkey", "Iran") ~ "Upper middle income",
    country %in% c("Kyrgyzstan", "Egypt", "Ivory Coast") ~ "Lower middle income",
    TRUE ~ income_group
  ))
```

```{r}
normal_behavior_income <- whr %>%
                   filter(year %in% 2015:2019) %>%
                   group_by(income_group) %>%
                   summarise(avg_score_pre_covid = mean(score, na.rm=TRUE))

abnormal_returns_income <- whr %>%
                    filter(year %in% 2020:2022) %>%
                    left_join(normal_behavior_income, by="income_group") %>%
                    mutate(abnormal_score = score - avg_score_pre_covid) %>%
                    select(country, year, income_group, score, avg_score_pre_covid, abnormal_score)

income_levels <- unique(abnormal_returns_income$income_group)

for (group in income_levels) {
  data_group <- subset(abnormal_returns_income, income_group == group)
  
  if (group == "Low income") {
    # Wilcoxon signed-rank test for 'Low income' group
    test_result <- wilcox.test(data_group$score, data_group$avg_score_pre_covid, paired=TRUE)
    cat(paste("Wilcoxon signed-rank test for income group:", group, "\n"))
  } else {
    # Paired t-test for other income groups
    test_result <- t.test(data_group$score, data_group$avg_score_pre_covid, paired=TRUE)
    cat(paste("T-test for income group:", group, "\n"))
  }
  
  print(test_result)
}


# Additional code to format output for LaTeX
for (group in income_levels) {
  data_group <- subset(abnormal_returns_income, income_group == group)
  
  if (group == "Low income") {
    # Wilcoxon signed-rank test for 'Low income' group
    test_result <- wilcox.test(data_group$score, data_group$avg_score_pre_covid, paired=TRUE)
    print(paste0("Wilcoxon signed-rank test for income group: ", group, "\\par"))
  } else {
    # Paired t-test for other income groups
    test_result <- t.test(data_group$score, data_group$avg_score_pre_covid, paired=TRUE)
    print(paste0("T-test for income group: ", group, "\\par"))
  }

  # Format the output for LaTeX
  cat("\\begin{table}[H]\n\\centering\n\\begin{tabular}{|l|c|}\n\\hline\n")
  cat(paste0("Test & ", ifelse(group == "Low income", "Wilcoxon signed-rank test", "Paired t-test"), "\\\\ \\hline\n"))
  cat(paste0("Group & ", group, "\\\\ \\hline\n"))
  cat(paste0("Statistic & ", test_result$statistic, "\\\\ \\hline\n"))
  cat(paste0("P-value & ", format.pval(test_result$p.value, digits = 2), "\\\\ \\hline\n"))
  cat(paste0("Confidence Interval & [", signif(test_result$conf.int[1], digits = 2), ", ", signif(test_result$conf.int[2], digits = 2), "]\\\\ \\hline\n"))
  cat("\\end{tabular}\n\\caption{Test result for ", group, "}\n\\end{table}\n\n")
}


```
# Proportional Contribution Analysis
## Visual Analysis

```{r}
whr <- whr %>%
  mutate(
    prop_gdp = gdp / score,
    prop_health = health / score,
    prop_social_support = social_support / score,
    prop_freedom = freedom / score,
    prop_corruption = corruption / score,
    prop_generosity = generosity / score
  )
```


```{r}
whr_long <- whr %>%
  select(year, prop_gdp, prop_health, prop_social_support, prop_freedom, prop_corruption, prop_generosity) %>%
  pivot_longer(-year, names_to="component", values_to="proportion")

whr_long %>%
  group_by(year, component) %>%
  summarise(mean_proportion = mean(proportion, na.rm = TRUE)) %>%
  ggplot(aes(x = year, y = mean_proportion, color = component)) +
  geom_line(size = 1) +
  theme_minimal() +
  labs(title = "Mean Proportional Contribution of Happiness Components Over Years",
       y = "Mean Proportional Contribution",
       x = "Year",
       color = "Component") +
  theme(legend.position = "bottom")
```

## Statistical Analysis
```{r warning=FALSE}
pre_years = c(2015, 2016, 2017, 2018, 2019)
post_years = c(2020, 2021, 2022)
components = c("prop_gdp", "prop_health", "prop_social_support", "prop_freedom", "prop_corruption", "prop_generosity")

# Summarize by average proportions for each period and rename columns accordingly
pre_summary <- whr %>%
  filter(year %in% pre_years) %>%
  group_by(country) %>%
  summarise(across(components, mean, na.rm=TRUE)) %>%
  rename_at(vars(components), ~ paste0("pre_", .))

post_summary <- whr %>%
  filter(year %in% post_years) %>%
  group_by(country) %>%
  summarise(across(components, mean, na.rm=TRUE)) %>%
  rename_at(vars(components), ~ paste0("post_", .))

# Merge the summaries
combined_data <- inner_join(pre_summary, post_summary, by="country")

# Initialize a results list
results <- list()

# Conduct paired t-tests for each component
for (comp in components) {
  pre_colname <- paste0("pre_", comp)
  post_colname <- paste0("post_", comp)
  
  test_result <- t.test(combined_data[[pre_colname]], combined_data[[post_colname]], paired=TRUE)
  results[[comp]] <- test_result
}

# View the results
results
```

# Regression Analysis

```{r}
whr <- whr %>%
  mutate(covid_ind = ifelse(year <= 2019, 0, 1)) %>%
  mutate(income_group = as.factor(income_group))

whr$centered_gdp_ppp <- whr$gdp_ppp - mean(whr$gdp_ppp, na.rm = TRUE)
whr$centered_life <- whr$life - mean(whr$life, na.rm = TRUE)
whr$centered_unemprate <- whr$unemp_rate - mean(whr$unemp_rate, na.rm = TRUE)
whr$centered_inflation <- whr$inflation - mean(whr$inflation, na.rm = TRUE)
whr$centered_population <- whr$population - mean(whr$population, na.rm = TRUE)
whr$centered_pcturbanpopulation <- whr$pct_urban_population - mean(whr$pct_urban_population, na.rm = TRUE)
```


## Year Fixed Effect 

```{r}
library(plm)
```

```{r}
fe_year <- plm(score ~ centered_gdp_ppp + centered_life + centered_unemprate + centered_population + centered_pcturbanpopulation + centered_inflation, 
                data = whr, 
                index = "year", 
                model = "within")
summary(fe_year)
```

## Year Fixed Effect + interaction
```{r}
fe_year_covid <- plm(score ~ centered_gdp_ppp + centered_life + centered_unemprate + centered_population + centered_pcturbanpopulation + centered_inflation + centered_gdp_ppp:covid_ind + centered_life:covid_ind + centered_unemprate:covid_ind + centered_pcturbanpopulation:covid_ind + centered_population:covid_ind + centered_inflation:covid_ind, 
                data = whr, 
                index = "year", 
                model = "within")

summary(fe_year_covid)

```

## Income Fixed Effect

```{r}
fe_income <- plm(score ~ centered_gdp_ppp + centered_life + centered_unemprate + centered_population + centered_pcturbanpopulation + centered_inflation + covid_ind, 
                data = whr, 
                index = "income_group", 
                model = "within")

summary(fe_income)
```

## Income Fixed Effect + interaction
```{r}
fe_income_covid <- plm(score ~ centered_gdp_ppp + centered_life + centered_unemprate + centered_population + centered_pcturbanpopulation + centered_inflation + covid_ind + centered_gdp_ppp:covid_ind + centered_life:covid_ind + centered_unemprate:covid_ind + centered_pcturbanpopulation:covid_ind + centered_population:covid_ind + centered_inflation:covid_ind, 
                data = whr, 
                index = "income_group", 
                model = "within")

summary(fe_income_covid)

```

## 2SLS
```{r}
library(AER)
whr$centered_medianage <- whr$median_age - mean(whr$median_age, na.rm = TRUE)
```


```{r}
first_stage <- lm(centered_gdp_ppp ~ centered_medianage + centered_life + centered_unemprate + centered_population + centered_pcturbanpopulation + centered_inflation + covid_ind, data = whr)

summary(first_stage)
```

```{r}
iv_model <- ivreg(score ~ centered_gdp_ppp + centered_life + centered_unemprate + centered_population + centered_pcturbanpopulation + centered_inflation + covid_ind | centered_medianage + centered_life + centered_unemprate + centered_population + centered_pcturbanpopulation + centered_inflation + covid_ind, data = whr)

summary(iv_model)

```

### Test

```{r}
library(lmtest)

coef_diff <- coef(iv_model) - coef(ols_centered)
diff_var <- vcov(iv_model) - vcov(ols_centered)

valid_coef <- names(coef_diff) %in% names(coef(iv_model))
coef_diff <- coef_diff[valid_coef]
diff_var <- diff_var[valid_coef, valid_coef]

hausman_stat <- t(coef_diff) %*% solve(diff_var) %*% coef_diff
hausman_stat <- as.numeric(hausman_stat)

df <- length(coef_diff)  
p_value <- pchisq(hausman_stat, df, lower.tail = FALSE)

list(statistic = hausman_stat, p_value = p_value)
```


